---
title: "Getting Started with healthcareai"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with healthcareai}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
set.seed(43170)
knitr::opts_chunk$set(echo = TRUE, results = "hold", collapse = TRUE, 
                      comment = "#>")
```

First we attach the healthcareai R package to make its functions available. If your package version is less than 2.0, none of the code here will work. You can check the package version with `packageVersion("healthcareai")`, and you can get the latest, cutting-edge development version with `install.packages("remotes"); remotes::install_github("HealthCatalyst/healthcareai-r")`.

```{r}
library(healthcareai)
```

`healthcareai` comes with a built in dataset documenting diabetes among adult Pima females. Once you attach the package, the dataset is available in the variable `pima_diabetes`. Let's take a look at the data with the `str` function. There are 768 records in 10 variables including one identifier column, several nominal variables, and substantial missingness (represented in R by `NA`).

```{r}
str(pima_diabetes)
```

# Quick Machine Learning

If you don't want to fuss with details any more than necessary, `machine_learn` is the function for you. It makes it as easy as possible to implement machine learning models by putting all the detais in the background so that you don't have to worry about them. Of course it might be wise to worry about them, and we'll get to how to do that further down, but for now, you can automatically take care of problems in the data, do basic feature engineering, and tune multiple machine learning models using cross validation with `machine_learn`.

`machine_learn` always gets the name of the data frame, then any columns that should not be used by the model (uninformative columns, such as IDs), then the variable to be predicted with `outcome =`.

```{r}
quick_models <- machine_learn(pima_diabetes, patient_id, outcome = diabetes)
```

`machine_learn` has told us that it has created a recipe for data preparation (this allows us to do exactly the same data cleaning and feature engineering when you want predictions on a new dataset), is ignoring `patient_id` when tuning models as we told it to, is training classification algorithms because the outcome variable `diabetes` is categorical, and has executed cross validation for two machine learning models: random forests, and k-nearest neighbors. Let's see what the models look like.

```{r}
quick_models
```

Everything looks as expected, and the best model is is a random forest that achives performance of AUROC = `r round(max(quick_models[[1]]$results$ROC), 2)`. Not bad for one line of code.

Now that we have our models, we can make predictions using the `predict` function. If you provide a new data frame to `predict` it will make predictions on the new data; otherwise, it will make predictions on the training data.

```{r}
predictions <- predict(quick_models)
predictions
```

We get a message about when the model was trained and how well it preformed in training, and we get back a data frame that looks sort of like the original, but has a new column `predited_diabetes` that contains the model-generated probability each individual has diabetes, and contains changes that were made preparing the data for model training, e.g. missingness has been filled in and `weight_class` has been split into a series of "dummy" variables.

We can plot how effectively the model is able to separate diabetic from non-diabetic individuals by simply calling the `plot` function on the output of `predict`.

```{r}
plot(predictions)
```

# Data Profiling

It is always a good idea to be aware of where there are missing values in data. The `missingness` function makes it easy to do that. In addition to looking for values R sees as missing, it looks for other values that might represent missing, such as `"NULL"`, and issues a warning if it finds any.

```{r}
missingness(pima_diabetes)
```

It's good that we don't have any missingness in our ID or outcome columns. We'll see how missingness in predictors is addressed further down.

# Data Preparation

To get an honest picture of how well a model performs (and an accurate estimate of how well it will perform on yet-unseen data), it is wise to hide a small portion of observations from model training and assess model performance on this "validation" or "test" dataset. In fact, `healthcareai` does this automatically and repeatedly under the hood, so it's not strictly necessary, but it's still a good idea. The `split_train_test` function makes this easy, and it ensures the test dataset has proportionally similar characteristics to the training dataset. By default, 80% of observations are used for training; that proportion can be adjusted with the `p` parameter. The `seed` parameter controls randomness so that you can get the same split every time you run the code if you want strict reproducability.

```{r}
split_data <- split_train_test(d = pima_diabetes,
                               outcome = diabetes,
                               p = .9,
                               seed = 84105)
```

`split_data` contains two data frames, named `train` and `test`.

One of the major workhorse functions in `healthcareai` is `prep_data`. It is called under-the-hood by `machine_learn`, so you don't have to worry about these details if you don't want to, but eventually you'll want to customize how your data is prepared; this is where you do that. The helpfile `?prep_data` describes what the function does and how it can be customized. Here, let's customize preparation to scale and center numeric variables and avoid collapsing rare factor levels into "other".

The first arguments to `prep_data` are the same as those to `machine_learn`: data frame, ignored columns, and the outcome column. Then we can specify prep details.

```{r}
prepped_training_data <- prep_data(split_data$train, patient_id, outcome = diabetes,
                                   center = TRUE, scale = TRUE,
                                   collapse_rare_factors = FALSE)
```

The "recipe" that the above message refers to is a set of instructions for how to transform a dataset the way we just transformed our training data. Any machine learning that we do (within `healthcareai`) on `prepped_training_data` will retain that recipe and apply it before making predictions on new data. That means that when you have models making predictions in production, you don't have to figure out how to transform the data or worry about encountering missing data or new category levels.

# Model Training

`machine_learn` takes care of data preparation and model training for you, but if you want more precise control, `tune_models` and `flash_models` are the model-training function you're looking for. They differ in that `tune_models` searches over hyperparameters to optimize model performance, while `flash_models` trains models at set hyperparameter values. So, `tune_models` produces better models, but takes longer (approaching 10x longer at default settings).

Let's just tune random forests (default also trains k-nearest neighbors), and to try to really optimize model performance, let's crank up `tune_depth`. That will tune the models over more combinations of hyperparameter values in the search for the best model.

Let's also select "PR" as our model metric. That optimizes for area under the precision-recall curve rather than the default of area under the receiver operating characteristic curve ("ROC"). This is usually a good idea when one outcome category is much more common than the other category.

```{r}
models <- tune_models(d = prepped_training_data,
                      outcome = diabetes,
                      models = "RF",
                      tune_depth = 50,
                      metric = "PR")
```

We get a message saying the training may take a while because we're training so many models, but in this case it takes just a couple minutes to train all those models.

We can examine how the model performs across hyperparameters by simply plotting the model object. It looks like extratrees is a superior split rule for this model, and larger values of minimum node size tend to do better.

```{r, fig.height = 7}
plot(models)
```

## Faster Model Training

Suppose our data actually consisted of millions of observations, rather than hundreds. It might take a very long time to train so many models, and while there's no guarantee that a model trained on all those observations will be optimal at the same hyperparameter values as a model trained on a subset of the observations, it could be useful to tune models on a smaller dataset and use the optimal settings on a larger dataset. You can do that by passing hyperparameter values to `flash_models`. If you don't pass hyperparameters to `flash_models` it will use sensible defaults; run `get_hyperparameter_defaults()` see the default values and get a list you can modify to customize.

Arguments to `flash_models` are otherwise very similar to `tune_models`.

```{r}
flash_models(d = prepped_training_data,
             outcome = diabetes,
             models = "RF",
             metric = "PR",
             hyperparameters = list(
               RF = list(
                 mtry = 10,
                 splitrule = "extratrees",
                 min.node.size = 20)))
```

# Prediction

We can easily get predictions on our original dataset. `predict` will automatically use the best-performing model from training (evaluated out-of-fold in cross validation). The predicted probabilities appear in the `predicted_diabetes` column.

```{r}
predict(models)
```

To get predictions on a new dataset, simply pass the new data to `predict`, and it will automatically be prepared based on the recipe generated on the training data. We can plot the predictions to see how well our model is doing, and we see that it's separating diabetic from non-diabetic individuals pretty well, although there a fair number of non-diabetics with high predicted probabilities of diabetes. This may be due to optimizing for precision recall, or may indicate pre-diabetic patients.

```{r}
test_predictions <- predict(models, split_data$test)
plot(test_predictions)
```

# A Regression Example

All the examples above have been classification tasks, predicting a yes/no outcome. Here's an example of a full regression modeling pipeline on a silly problem: predicting individuals' ages. The code is very similar to classification.

```{r}
regression_models <- machine_learn(pima_diabetes, patient_id, outcome = age)
summary(regression_models)
```

Let's make a prediction on a hypothetical new patient. Note that the model handles missingness in `insulin` and a new category level in `weight_class` without a problem (but warns about it).

```{r}
new_patient <- data.frame(
  pregnancies = 0,
  plasma_glucose = 80,
  diastolic_bp = 55,
  skinfold = 24,
  insulin = NA,
  weight_class = "???",
  pedigree = .2,
  diabetes = "N")
predict(regression_models, new_patient)
```

